|

### 2021-07-27 article/articleCustomerDropoutMembership.Rmd
upd

---
title: |
  |
  |
  |
  | \vspace{1cm}Customer dropout membership^[Corresponding address: sobreiro@esdrm.ipsantarem.pt. The current template adapts part of the Rmd code by [Paul C. Bauer](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown), Mannheim Centre for European Social Research.]\vspace{0.5cm}
author: |
  | Pedro Sobreiro, Javier Berrocal, Domingos Martinho, José Garcia Alonso
  |
  | Extremadura University
  |
date: |
  |
  |
  | `r gsub("^0", "", format(Sys.time(), "%d %B, %Y"))`
  |
  |
linestretch: 1.2
colorlinks: true
abstract: \noindent\setstretch{1}Abstract of the article. Here we can place more info.\vspace{.8cm}
bibliography: refMembership.bib
csl: american-sociological-association.csl
output:
  bookdown::pdf_document2:
    includes:
    toc: no
    keep_tex: true
    latex_engine: xelatex
    number_sections: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin = 1in
always_allow_html: yes
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{color}
   - \usepackage{pdfpages}
   - \usepackage{amsmath}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{hyperref}
editor_options: 
  markdown: 
    wrap: 80
---

@@

@@

### Data description

@@

names(df_members)
# rename column labels
names(df_members) <- c("num_socio", "dt_inscription", "year", "birth_date",
                       "age", "sex", "marital_status", "category",
                       "monthly_fee", "occupation", "zip_code",
                       "dt_last_invoice", "dt_last_payment", "total_amount",
                       "total_matches", "season_matches",
                       "days_since_last_payment", "months_since_last_payment",
                       "dropout", "years_membership", "stadium_access",
                       "quart_stadium_entries", "inscription_month")

@@

# select relevant variables
df_members <- df_members %>%
select(year, age, sex, marital_status, monthly_fee, total_amount, total_matches,
       season_matches, months_since_last_payment, dropout, years_membership,
       stadium_access, quart_stadium_entries,inscription_month)

@@

Table \@ref(tab:summarytable) shows data's summary statistics. 
The average age is `r round(mean(df_members$age,na.rm = TRUE),1)` ±
`r round(sd(df_members$age,na.rm = TRUE),1)`, the members have an attendance 
`r round(mean(df_members$total_matches,na.rm = TRUE),0)` ± `r 
round(sd(df_members$total_matches,na.rm = TRUE),1)` with a membership
`r round(mean(df_members$years_membership,na.rm = TRUE),0)` ± `r 
round(sd(df_members$years_membership,na.rm = TRUE),1)`.

```{r teste, eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE, paged.print=FALSE}
library(stargazer)
df_summary <- summary(df_members)
stargazer(df_members,
          title = "Summary statistics",
          label = "tab1cars",
          table.placement = "h",
          header = FALSE,
          summary = TRUE,
          summary.stat = c("min", "p25", "median", "p75", "max", "median", "sd")
         )
```
```{r summarytable, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(gtsummary)
library(kableExtra)
library(labelled)

var_label(df_members$year) <- "Inscription year"
var_label(df_members$age) <- "Age in years"
var_label(df_members$sex) <- "Male or female"
var_label(df_members$marital_status) = "Single, married and other."

tbl <- df_members %>%
    tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})",
                                 all_categorical() ~ "{p}%"),
                type = list(age ~ "continuous")
               ) %>% add_stat_label()

as_kable_extra(tbl,booktabs=T,caption="Summary statistics of features used")
```
Figure \@ref(fig:membershipyear) shows the distribution of the dropout considering the inscription year.

```{r membershipyear, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Number of members by year"}
members_year<-df_members%>% 
  select(year,dropout) %>%
  group_by(year,dropout) %>%
  summarize(count = n())

members_year$dropout = factor(members_year$dropout)  
# scale_y_continuous(breaks=c(2,4,6,8,10))+
g<-ggplot(data = members_year,mapping = aes(x = year,y = count,linetype = dropout)) 
g+geom_line()+labs(x='year',y='Number of members') +
  theme_classic()
```

## TODO

@@

## Running the model

    from pysurvival.utils.display import correlation_matrix
    correlation_matrix(df[features], figure_size=(10,10), text_fontsize=8)

@@

correlation_matrix(df_members, figure_size=(10,10), text_fontsize=6)

r.df_members = df_members

```

Removed the variables with greater correlations

    to_remove = ['totalJogos', 'idaEstadio']
    features = np.setdiff1d(features, to_remove).tolist()

## Model building

@@

    from pysurvival.models.survival_forest import RandomSurvivalForestModel
    csf = RandomSurvivalForestModel(num_trees=200)
    csf.fit(X_train, T_train, E_train, max_features='sqrt', max_depth=5, min_node_size=20)

The model accuracy is very high in the first years. The prediction is very
similar to the actual value.

![Prediction accuracy](../analysis/predictionAccuracy.png)

All the outputs are available [here](../analysis/01.survRandomForest.pdf)

@@

```{=latex}
\begin{table}[ht]
    \footnotesize
    \centering
    \begin{tabular}{p{0.15\textwidth}p{0.55\textwidth}p{0.25\textwidth}}
    \hline
    \textbf{Model name} & \textbf{Explored parameters values} & \textbf{Best parameters} \\
    \hline
    Survival trees  & pysurvival random forest                  & a \\
    Survival trees
    with clusters   & pysurvival random forest with clusters    & a \\
    Scikit survival 
    trees           & scikit survival                           & a \\
    Scikit survival 
    with clusters   & scikit with clusters                      & a \\
    Scikit survival 
    gradient boost  & scikit survival gradient boost            & a \\
    Scikit survival 
    gradient boost 
    with clusters   & scikit with clusters                      & a \\
    \hline
    \end{tabular}
    \caption{Hyper-parameters best values}
    \label{hyperparametersbestvalues}
\end{table}
```
```{=html}
<!--
|        Model name        | Explored hyper-parameters' values                    | Best parameters |
| ------------------------ | ---------------------------------------------------- | --------------- |
| Survival trees           |                                                      |                 |   
| without cluster          |                                                      |                 |  
-->
```

# Conclusion

>>>>

---
title: |
  |
  |
  |
  | \vspace{1cm}Customer dropout membership^[Corresponding address: sobreiro@esdrm.ipsantarem.pt. Quality of Life Research Centre, Polytechnic Institute of Santarém, Portugal]\vspace{0.5cm}
author: |
  | Pedro Sobreiro, Javier Berrocal, Domingos Martinho, José Garcia Alonso
  |
  | Extremadura University
  |
date: |
  |
  |
  | `r gsub("^0", "", format(Sys.time(), "%d %B, %Y"))`
  |
  |
linestretch: 1.2
colorlinks: true
abstract: \noindent\setstretch{1}Abstract of the article. Here we can place more info.\vspace{.8cm}
bibliography: refMembership.bib
csl: american-sociological-association.csl
output:
  bookdown::pdf_document2:
    includes:
    toc: no
    keep_tex: true
    latex_engine: xelatex
    number_sections: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin = 1in
always_allow_html: yes
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{color}
   - \usepackage{pdfpages}
   - \usepackage{amsmath}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{hyperref}
editor_options: 
  markdown: 
    wrap: 80
---

@@

The model performance was determined with the concordance probability (C-index),
Brier Score (BS) and Mean Absolute Error (MAE) [@wangmachine2017]. The feature
importance was determined calculating the difference between the true class
label and noised data [@Breiman_2001].

@@

@@

names(df_members)
# rename column labels
names(df_members) <- c("num_socio", "dt_inscription", "year_inscription", "birth_date",
                       "age", "sex", "marital_status", "category",
                       "monthly_fee", "occupation", "zip_code",
                       "dt_last_invoice", "dt_last_payment", "total_amount",
                       "total_matches", "season_matches",
                       "days_since_last_payment", "months_since_last_payment",
                       "dropout", "years_membership", "stadium_access",
                       "quart_stadium_entries", "inscription_month")

@@

# select relevant variables
df_members <- df_members %>%
select(age, sex, marital_status, monthly_fee, total_amount, total_matches,
       season_matches, months_since_last_payment, dropout, years_membership,
       stadium_access, quart_stadium_entries,inscription_month)

@@

Table \@ref(tab:summarytable) shows data's summary statistics. The average age
is `r round(mean(df_members$age,na.rm = TRUE),1)` ±
`r round(sd(df_members$age,na.rm = TRUE),1)`, the members have an attendance of
`r round(mean(df_members$total_matches,na.rm = TRUE),0)` ±
`r  round(sd(df_members$total_matches,na.rm = TRUE),1)` with a membership of
`r round(mean(df_members$years_membership,na.rm = TRUE),0)` ±
`r  round(sd(df_members$years_membership,na.rm = TRUE),1)` years.

```{r teste, eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE, paged.print=FALSE}
library(stargazer)
df_summary <- summary(df_members)
stargazer(df_members,
          title = "Summary statistics",
          label = "tab1cars",
          table.placement = "h",
          header = FALSE,
          summary = TRUE,
          summary.stat = c("min", "p25", "median", "p75", "max", "median", "sd")
         )
```

```{r summarytable, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(gtsummary)
library(kableExtra)
library(labelled)

var_label(df_members$age) <- "Age in years"
var_label(df_members$sex) <- "Male or female"
var_label(df_members$marital_status) = "Single, married and other."

tbl <- df_members %>%
    tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})",
                                 all_categorical() ~ "{p}%"),
                type = list(age ~ "continuous")) %>%
    add_stat_label()

as_kable_extra(tbl, booktabs = T,
               caption = "Summary statistics of features used")
```

Figure \@ref(fig:membershipyear) shows the distribution of the dropout
considering the number of years of membership.

```{r membershipyear, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Number of members by year"}
members_year <- df_members %>%
  select(years_membership, dropout) %>%
  group_by(years_membership, dropout) %>%
  summarize(count = n())

members_year$dropout <- factor(members_year$dropout)
g <- ggplot(data = members_year,
            mapping = aes(x = years_membership, y = count,linetype = dropout))
g + geom_line() + labs(x = "year", y = "Number of members") +
    theme_classic()
```

## Model construction

Address the model construction... the categorical variables $sex$,
$marital\_status$ and $quart\_stadium\_entries$ where converted to dummy
variables.

@@

### Survival trees based model

In this model... The survival trees based model uses pysurvival random forest

@@

correlation_matrix(df_members[features], figure_size=(10,10), text_fontsize=6)

r.df_members = df_members
```

Removed the variables with greater correlations $total\_matches$ and
$quart\_stadium\_entries$

```{python removeVar, eval=TRUE, echo=FALSE, warning=FALSE}
to_remove = ['total_matches', 'quart_stadium_entries_ate 1']

#consider also the features previously removed
features = np.setdiff1d(df_members[features].columns, to_remove).tolist()

df_members.drop(columns = to_remove, inplace=True)

correlation_matrix(df_members[features], figure_size=(10,10), text_fontsize=6)
```

```{python createModel, eval=TRUE, echo=FALSE, cache = TRUE, warning = FALSE}
from pysurvival.models.survival_forest import RandomSurvivalForestModel

# Number of samples in the dataset
N = df_members.shape[0]

# Building training and testing sets
from sklearn.model_selection import train_test_split
index_train, index_test = train_test_split( range(N), test_size = 0.4)
data_train = df_members.loc[index_train].reset_index( drop = True )
data_test  = df_members.loc[index_test].reset_index( drop = True )

# Creating the X, T and E inputs
X_train, X_test = df_members[features], data_test[features]
T_train, T_test = df_members[time_column], data_test[time_column]
E_train, E_test = df_members[event_column], data_test[event_column]

# Fitting the model
csf = RandomSurvivalForestModel(num_trees=200)
csf.fit(X_train, T_train, E_train, max_features='sqrt',
        max_depth=5, min_node_size=20)
```

Table \@ref(tab:summarytable2) shows variables importance.

```{r summarytable2, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
tbl <- py$csf$variable_importance_table
kbl(tbl, booktabs = T, caption = "Summary statistics of features used")
```

#### Model building

@@

```{python pysurvivalperformance, eval=TRUE, echo=FALSE, cache=TRUE, warning = FALSE}
from pysurvival.utils.metrics import concordance_index
from pysurvival.utils.display import integrated_brier_score
from pysurvival.utils.display import compare_to_actual

c_index = concordance_index(csf, X_test, T_test, E_test)

ibs = integrated_brier_score(csf, X_test, T_test, E_test, t_max=12,
    figure_size=(12,5))

results = compare_to_actual(csf, X_test, T_test, E_test, is_at_risk = False,  
figure_size=(12, 6), metrics = ['rmse', 'mean', 'median'])
```

The model accuracy is very high in the first years. The prediction is very
similar to the actual value. The absolute error mean of \~37 customers.

### Survival trees based model with clusters

Here we are will create clusters and developed the optimization within each
cluster...

The calculation of he number of clusters used the package mclust
[@scrucca2016] using the Bayesian Information Criterion (BIC). The model that
gives the minimum BIC score can be selected as the best model [@schwarz1978]
simplifying the problem related to choosing the number of components and
identifying the structure of the covariance matrix, based on modelling with
multivariate normal distributions for each component that forms the data set
[@akogul2016].

In multivariate models are available the following approaches:

-   "EII"spherical, equal volume
-   "EEE"ellipsoidal, equal volume, shape, and orientation
-   "VII"spherical, unequal volume
-   "VVV"ellipsoidal, varying volume, shape, and orientation, which is used as
    default for initialization of EM algorithm
-   VVI": diagonal, varying volume and shape t

estava a aqui a confirmar os clusters, talvez seja melhor reduzir as variáveis...

```{r clusters, echo=FALSE, message=FALSE, warning=FALSE}
library(mclust)
y <- scale(py$df_members)
bic <- mclustBIC(y,modelNames = c("VEI"))
plot(bic)

temp = data.frame(bic[,'VEI'])
temp$cluster = row.names(temp)

names(temp) = c("val","cluster")

kbl(temp, booktabs = T, caption = "Summary statistics of features used")

mod1 <- Mclust(py$df_members,x=BIC)
summary(mod1)

#plot(mod1, what="uncertainty")
```

TODO:

@@

```{=latex}
\begin{table}[ht]
    \footnotesize
    \centering
    \begin{tabular}{p{0.15\textwidth}p{0.55\textwidth}p{0.25\textwidth}}
    \hline
    \textbf{Model name} & \textbf{Explored parameters values} & \textbf{Best parameters} \\
    \hline
    Survival trees  & pysurvival random forest                  & a \\
    Survival trees
    with clusters   & pysurvival random forest with clusters    & a \\
    Scikit survival 
    trees           & scikit survival                           & a \\
    Scikit survival 
    with clusters   & scikit with clusters                      & a \\
    Scikit survival 
    gradient boost  & scikit survival gradient boost            & a \\
    Scikit survival 
    gradient boost 
    with clusters   & scikit with clusters                      & a \\
    \hline
    \end{tabular}
    \caption{Hyper-parameters best values}
    \label{hyperparametersbestvalues}
\end{table}
```
```{=html}
<!--
|        Model name        | Explored hyper-parameters' values                    | Best parameters |
| ------------------------ | ---------------------------------------------------- | --------------- |
| Survival trees           |                                                      |                 |   
| without cluster          |                                                      |                 |  
-->
```
# Conclusion
